@comment -*-texinfo-*-
@comment $Id: copyright.tex,v 1.15 1998-06-18 05:31:04 obachman Exp $
@comment this file contains the copyright notice on Singular

@c The following directives are necessary for proper compilation
@c with emacs (C-c C-e C-r).  Please keep it as it is.  Since it
@c is wrapped in `@ignore' and `@end ignore' it does not harm `tex' or
@c `makeinfo' but is a great help in editing this file (emacs
@c ignores the `@ignore').
@ignore
%**start
\input texinfo.tex
@setfilename copyright.info
@node Top, Preface
@menu
* Preface::
@end menu
@node Preface, General concepts, Top, Top
@chapter Preface
%**end
@end ignore

@cindex copyright
@cindex factory
@cindex GMP
@cindex MP
@cindex readline
@cindex libfac

@center @sc{Singular} version @value{VERSION}
@center University of Kaiserslautern
@center Department of Mathematics and Centre for Computer Algebra
Authors: G.-M. Greuel, G. Pfister, H.
@tex
Sch\"onemann
@end tex
@ifinfo
Schoenemann
@end ifinfo

Copyright @copyright{} 1986-98; All Rights reserved
@sp 2
                                @center NOTICE

Permission to use, copy, modify, and distribute this software or parts
thereof and its documentation or parts thereof for non-commercial
purposes and without fee is hereby granted provided the following four
conditions are satisfied:

@enumerate
@item
The above copyright notice appears in all copies of the software
and both the copyright notice and this permission notice
appear in supporting documentation.

@item
The name of the program ("@sc{Singular}") is retained.

@item
Portions of the software which are modified are marked.

@item 
You have registered yourself as a Singular user by sending email to
@email{singular@@mathematik.uni-kl.de} with the subject line (or, mail body)
@code{register}. 
@end enumerate

Neither the University of Kaiserslautern nor the authors make any
representations about the suitability of this software for any
purpose.  This software is provided ``as is'' without express or
implied warranty.

If your intended use of @sc{Singular} is not covered by the license above,
please contact us.  Notice that in the license above we have not
granted permission to make copies of Singular to be sold, distributed
on media which are sold, or distributed along with software which is
sold. 

If you use Singular or parts thereof in a project and/or publish
results that were partly obtained using @sc{Singular}, we request that you
cite @sc{Singular} (see
@url{http://www.ma@-the@-ma@-tik.uni-kl.de/~zca/Singular/how_to_cite.html} 
on information on how to cite Singular) and inform us about it.

Please send any comments or bug reports to
@email{singular@@mathematik.uni-kl.de}.

The following parts of @sc{Singular} have their own copyright: the Gnu Multiple
Precision Library (GMP), the Multi Protocol library (MP), the Readline library,
the Factory library, and the libfac library.
Their copyrights and licences can be found in the accompanying files
which are distributed along with these libraries.

@section Availability

The latest information about @sc{Singular} is always available from
@url{http://www.ma@-the@-ma@-tik.uni-kl.de/~zca/Singular}.
The program @sc{Singular} and the above mentioned parts are available via
anonymous ftp through the following addresses:
@table @asis
@item GMP
@copyright{} Free Software Foundation:
@code{libgmp-2.0.2.tar.gz}
@* @url{ftp://ftp.gnu.ai.mit.edu} or its mirrors
@item readline
@copyright{} Free Software Foundation:
@code{libreadline-2.0.tar.gz}
@* @url{ftp://ftp.gnu.ai.mit.edu} or its mirrors
@item MP
@copyright{} Gray/Kajler/Wang, Kent State University:
@code{MP-1.1.2.tar.gz},
@* @url{http://mcs.kent.edu/pub/MP}
@item Factory
@copyright{} Greuel/Stobbe, University of Kaiserslautern:
@code{factory-1.3b.tar.gz}
@* Contact @email{factory@@mathematik.uni-kl.de} for details on Factory.
@item libfac
@copyright{}  Messollen, University of 
@tex
Saarbr\"ucken:
@end tex
@ifinfo
Saarbruecken:
@end ifinfo
@code{libfac-0.3.0.tar.gz}
@* Contact @email{michael@@math.uni-sb.de} for details on libfac.
@item @sc{Singular} binaries
@url{ftp://www.mathematik.uni-kl.de/pub/Math/Singular/} or via a
WWW brow@-ser from
@url{http://www.mathematik.uni-kl.de/ftp/pub/Math/@-Singular/}
@end table

@ifset singularmanual
@section Acknowledgements

The development of @sc{Singular} is directed and coordinated by
Gert-Martin Greuel, Gerhard Pfister, and Hans
@tex
Sch\"onemann.
@end tex
@ifinfo
Schoenemann.
@end ifinfo

Contributions to the kernel of @sc{Singular} have been made by
Olaf Bachmann,
Hubert Grassmann,
@tex
Kai Kr\"uger,
@end tex
@ifinfo
Kai Krueger,
@end ifinfo
Wolfgang Neumann,
@tex
Thomas N\"u{\ss}ler,
@end tex
@ifinfo
Thomas Nuessler,
@end ifinfo
Wilfred Pohl,
Jens Schmidt,
Thomas Siebert,
@tex
R\"udiger Stobbe,
@end tex
@ifinfo
Ruediger Stobbe,
@end ifinfo
and
Tim Wichmann.

@sc{Singular} libraries have furthermore been contributed by 
Wolfram Decker,
Christian Gorzel,
Agnes Eileen Heydtmann,
Ulrich Klein,
Martin Lamm,
and
Bernd Martin.

We should like to acknowledge the financial support given by
the Volkswagen-Stiftung,
the Deutsche Forschungsgemeinschaft
and the Stiftung
@ifinfo
fuer
@end ifinfo
@tex
f\"ur
@end tex
Innovation des Landes Rhein@-land-Pfalz
to the
@sc{Singular} project.

@end ifset
